
```{r read_cpi_and_unemp,echo=FALSE,fig.width=7,fig.height=2.5,out.width="5in",}
cpi_wide <- read.table(file="data/consumer_price_index.csv",sep=",",header=TRUE)
cpi_long <- as.vector(t(as.matrix(cpi_wide[,2:13])))
inflation <- diff(log(cpi_long))
year <- seq(from=1913,length=length(cpi_long),by=1/12)[-1]
inflation <- inflation[year>=1980]*12*100 # percent, annualized
year <- year[year>=1980]
i_smo <- loess(inflation~year,span=0.1)$fitted
inflation <- i_smo

unemp_wide <- read.table(file="data/adjusted_unemployment.csv",sep=",",header=TRUE)
unemployment <- as.vector(t(as.matrix(unemp_wide[,2:13])))
u_year <- seq(from=1948,length=length(unemployment),by=1/12)[-1]
unemployment <- unemployment[u_year>=1980]
u_year <- u_year[u_year>=1980]
ols <- lm(inflation~unemployment)

par(mai=c(0.5,1,0.35,0.1))
par(mfrow=c(1,2))
matplot(x=year,y=cbind(inflation,unemployment),ty="l",ylab="")
text("A",x=1988,y=13,cex=2)
acf(resid(lm(inflation~unemployment)),lag.max=50)
text("B",x=12,y=0.87,cex=2)

```

\(A) Inflation (black) and unemployment (red) for the USA, 1980-2024. (B) Sample autocorrelation function of the residuals from a least square regression, 'lm(inflation~unemployment)', with estimated coefficients below. Which is the best interpretation of these graphs and fitted model?

```{r lm_inflation_unemployment,echo=FALSE}
#summary(lm(inflation~unemployment))
lm(inflation~unemployment)$coef
```

A: 0.05 is a reasonable estimate for the additional unemployment caused by one percentage point of additional inflation. We should not trust the uncertainty estimate (not shown), since our model does not allow for autocorrelation of the residuals.

B: 0.05 is a reasonable estimate for the association between inflation and unemployment. We should not assume there is a causal relationship. We should not trust the uncertainty estimate (not shown), since our model does not allow for autocorrelation of the residuals.

C: 0.05 is a reasonable estimate for the association between inflation and unemployment.  We should make an additional assumption that there are no confounding variables, and then we can interpret this association to be causal. We should not trust the uncertainty estimate (not shown), since our model does not allow for autocorrelation of the residuals.



 



