---
title: "Quiz 1, STATS/DATASCI 531/631 W25	"
author: "In class on 2/17, 2:30pm to 3:00pm"
output:
  pdf_document:
    latex_engine: xelatex
  toc: no
header-includes:
   - \usepackage{ifthen}
csl: ecology.csl
---


```{r preliminaries,echo=FALSE,cache=FALSE}
ALL <- TRUE
SOL <- TRUE
#SOL <- FALSE
QLABELS <- TRUE
EXAM <- FALSE
#EXAM <- TRUE
if(EXAM){
  ALL <- FALSE
  SOL <- FALSE
  QLABELS <- FALSE
}
```

\newcommand\exam[1]{#1}
<!-- \renewcommand\exam[1]{} to remove exam formatting -->
\renewcommand\exam[1]{}
 
```{r setup,echo=FALSE,cache=FALSE}
# library(mat2tex)
library(png)
set.seed(48)
knitr::opts_chunk$set(cache = FALSE)

q_setup <- function(n){
  q_dirs <- list.files(pattern=paste0("^Q",n,"-"))
  cat_file <- paste0("tmp",n,".Rmd")
  #
  # first add lines so each question knows its directory
  #
  for(q_dir in q_dirs){
    q_file <- paste(q_dir,"/q.Rmd",sep="") 
    q_tmp <- paste(q_dir,"/tmp.Rmd",sep="") 
    file.create(q_tmp)  
    write(c(
      "```{r,echo=F,results=ifelse(QLABELS,'asis','hide')}",
      paste("my_dir <- '",q_dir,"'",sep=""),
      "cat('**',my_dir,'.**',sep='')",
      "```"),
      file=q_tmp,append=TRUE
    )
    file.append(q_tmp,q_file)
    write(c(
      "```{r,child=paste(my_dir,'/sol.Rmd',sep=''),eval=SOL}",
      "``` \n"
    ),file=q_tmp,append=TRUE)
  }
  # 
  # Then concatenate these enhanced question files
  #
  q_files <- paste(q_dirs,"/tmp.Rmd",sep="") 
  if(!ALL) q_files <- sample(q_files,size=1)
  file.create(cat_file)
  file.append(cat_file,q_files)
  cat_file
}

```



\newcommand\var{\mathrm{Var}}
\newcommand\varSample{\mathrm{var}}
\newcommand\cov{\mathrm{Cov}}
\newcommand\covSample{\mathrm{cov}}
\newcommand\cor{\mathrm{Cor}}
\newcommand\corSample{\mathrm{cor}}

\exam{\vspace{1.5cm}

{\bf Name: \hspace{6cm} UMID: }

\vspace{1.5cm}

}

```{r,results="asis",echo=F,purl=F,eval=ALL}
cat("**This document produces different random quizzes each time the source code generating it is run. The actual quiz will be a realization generated by this random process, or something similar.**")
```

```{r,results="asis",echo=F,purl=F,eval=ALL}
cat("**This version lists all the questions currently in the quiz generator. The actual quiz will have one question sampled from each of the 6 question categories.**")
```

**Instructions. You have a time allowance of 30 minutes. The quiz may be ended early if everyone is done. The quiz is closed book, and you are not allowed access to any notes. Any electronic devices in your possession must be turned off and remain in a bag on the floor.**

**For each question, circle one letter answer and provide some supporting reasoning.**


\exam{\newpage}

### Q1. Stationarity and unit roots.

```{r q1-setup,include=F}
cat_file1 <- q_setup(1)
```

```{r q1, child = cat_file1}
```

\exam{\newpage}

### Q2. Calculations for ARMA models

```{r q2-setup,include=F}
cat_file2 <- q_setup(2)
```

```{r q2, child = cat_file2}
```

\exam{\newpage}


### Q3. Likelihood-based inference for ARMA models 

```{r q3-setup,include=F}
cat_file3 <- q_setup(3)
```

```{r q3, child = cat_file3}
```

\exam{\newpage}

### Q4. Interpreting diagnostics

```{r q4-setup,include=F}
cat_file4 <- q_setup(4)
```

```{r q4, child = cat_file4}
```

\exam{\newpage}

### Q5. The frequency domain

```{r q5-setup,include=F}
cat_file5 <- q_setup(5)
```

```{r q5, child = cat_file5}
```

\exam{\newpage}

### Q6. Scholarship for time series projects

```{r q6-setup,include=F}
cat_file6 <- q_setup(6)
```

```{r q6, child = cat_file6}
```



\exam{\vfill}

------



License: This material is provided under a [Creative Commons license](https://ionides.github.io/531w25/LICENSE)<br>




------
