---
title: "Quiz 2, STATS/DATASCI 531/631 W25	"
author: "In class on 4/16, 2:30pm to 3:00pm"
output:
  pdf_document:
    latex_engine: xelatex
  toc: no
header-includes:
   - \usepackage{ifthen}
csl: ecology.csl
---


```{r preliminaries,echo=FALSE,cache=FALSE}
ALL <- TRUE
SOL <- TRUE
# SOL <- FALSE
QLABELS <- TRUE
EXAM <- FALSE
# EXAM <- TRUE
if(EXAM){
  ALL <- FALSE
  SOL <- FALSE
  QLABELS <- FALSE
}
```

\newcommand\exam[1]{#1}
<!-- uncomment to remove exam formatting
\renewcommand\exam[1]{}
-->

 
```{r setup,echo=FALSE,cache=FALSE}
# library(mat2tex)
library(png)
# set.seed(42)
knitr::opts_chunk$set(cache = FALSE)

q_setup <- function(n){
  q_dirs <- list.files(pattern=paste0("^Q",n))
  cat_file <- paste0("tmp",n,".Rmd")
  #
  # first add lines so each question knows its directory
  #
  for(q_dir in q_dirs){
    q_file <- paste(q_dir,"/q.Rmd",sep="") 
    q_tmp <- paste(q_dir,"/tmp.Rmd",sep="") 
    file.create(q_tmp)  
    write(c(
      "```{r,echo=F,results=ifelse(QLABELS,'asis','hide')}",
      paste("my_dir <- '",q_dir,"'",sep=""),
      "cat('**',my_dir,'.**',sep='')",
      "```"),
      file=q_tmp,append=TRUE
    )
    file.append(q_tmp,q_file)
    write(c(
      "```{r,child=paste(my_dir,'/sol.Rmd',sep=''),eval=SOL}",
      "``` \n"
    ),file=q_tmp,append=TRUE)
  }
  # 
  # Then concatenate these enhanced question files
  #
  q_files <- paste(q_dirs,"/tmp.Rmd",sep="") 
  if(!ALL) q_files <- sample(q_files,size=1)
  file.create(cat_file)
  file.append(cat_file,q_files)
  cat_file
}

```



\newcommand\var{\mathrm{Var}}
\newcommand\varSample{\mathrm{var}}
\newcommand\cov{\mathrm{Cov}}
\newcommand\covSample{\mathrm{cov}}
\newcommand\cor{\mathrm{Cor}}
\newcommand\corSample{\mathrm{cor}}

\exam{\vspace{1.5cm}

{\bf Name: \hspace{6cm} UMID: }

\vspace{1.5cm}

}

```{r,results="asis",echo=F,purl=F,eval=ALL}
cat("**This document produces different random quizzes each time the source code generating it is run. The actual quiz will be a realization generated by this random process, or something similar.**")
```

```{r,results="asis",echo=F,purl=F,eval=ALL}
cat("**This version lists all the questions currently in the quiz generator. The actual quiz will have one question sampled from each of the 6 question categories.**")
```

**Instructions. You have a time allowance of 30 minutes. The quiz may be ended early if everyone is done. The quiz is closed book, and you are not allowed access to any notes. Any electronic devices in your possession must be turned off and remain in a bag on the floor.**

**For each question, circle one letter answer and provide supporting reasoning. If you run out of space, you may continue on the back of the page, but please indicate to the reader that you are doing so.**


\exam{\newpage}

### Q1. Foundations of POMP models

```{r q1-setup,include=F}
cat_file1 <- q_setup(8)
```

```{r q1, child = cat_file1}
```
\exam{Supporting reasoning:}

\exam{\newpage}

### Q2. Likelihood evaluation; the particle filter

```{r q2-setup,include=F}
cat_file2 <- q_setup(9)
```

```{r q2, child = cat_file2}
```
\exam{Supporting reasoning:}

\exam{\newpage}


### Q3. Likelihood maximization; iterated filtering

```{r q3-setup,include=F}
cat_file3 <- q_setup(10)
```

```{r q3, child = cat_file3}
```
\exam{Supporting reasoning:}

\exam{\newpage}

### Q4. Data analysis: epidemiological models

```{r q4-setup,include=F}
cat_file4 <- q_setup(11)
```

```{r q4, child = cat_file4}
```
\exam{Supporting reasoning:}

\exam{\newpage}

### Q5. Data analysis: financial models

```{r q5-setup,include=F}
cat_file5 <- q_setup(12)
```

```{r q5, child = cat_file5}
```
\exam{Supporting reasoning:}

\exam{\newpage}

### Q6. Computing with POMP models

```{r q6-setup,include=F}
cat_file6 <- q_setup(13)
```

```{r q6, child = cat_file6}
```
\exam{Supporting reasoning:}



\exam{\vfill}

------



License: This material is provided under a [Creative Commons license](https://ionides.github.io/531w25/LICENSE)<br>




------
