
```{r nyc_injuries_timeplot,echo=FALSE,fig.width=5,fig.height=2.5,out.width="4in",warning=FALSE,message=FALSE}
# code adapted from
# https://ionides.github.io/531w24/midterm_project/project01/blinded.html
library(tidyverse)
Collision_data <- read.csv("data/nyc_motor_collision_injuries.csv",skip=4)
Collision_data$date <- as.Date(Collision_data$date, format="%m/%d/%Y")
Collision_data <- Collision_data %>% filter(date >= as.Date("2013-01-01")) %>% filter(date < as.Date("2024-01-01"))

monthly_data <- Collision_data %>%
  group_by(year = year(date), month = month(date)) %>%
  summarise(person_injured = sum(person_injured)) %>%
  ungroup() %>%
  mutate(date = as.Date(paste(year, month, "01", sep = "-"))) %>%
  select(date, person_injured)

ggplot(monthly_data, aes(x = date, y = person_injured/1000)) +
  geom_line() +
  labs(title = "", x = "Year", y = "injuries (thousands)") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") 
```

Above are monthly injuries from motor vehicle collisions in New York City. An augmented Dickey-Fuller test, `tseries::adf.test(injuries)`, gives a p-value of 0.01. Which is the best way to proceed:

```{r adf-for-nyc,echo=FALSE,warning=FALSE, message=FALSE,results=FALSE}
my_adf_print <- function(x) cat(
  "data:  ", x$data.name,
  "\nDickey-Fuller = ", x$statistics,
  ", Lag order = ", x$parameter,
  ", p-value = ", x$p.value,
  "\nAlternative hypothesis: ",x$alternative, "\n "
)
my_adf_print(tseries::adf.test(monthly_data$person_injured))
```

A: The time plot indicates a non-constant mean function describing a major dip due to the COVID-19 pandemic and an increasing trend at other times. The ADF test does not support or refute that model.

B: The ADF test suggests the series is stationary, supporting a decision to fit a SARMA model.

C: The ADF test suggests the series is non-stationary; it should be differenced before fitting a SARMA.

D: The ADF test indicates that the series is non-stationary, supporting the use of a non-constant mean function to describe a major dip due to the COVID-19 pandemic and an increasing trend at other times.


